\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
Wrote the rest of code region subsubsection.

\subsubsection{Code region}
The code region is a central region to the logical relation used to model the code memory of a component.
It is especially interesting because it specifies what the return and closure seals in a component can be used to seal.
The code region is defined as follows:
\[
  \codereg{\sigrets,\sigcloss,\mscode,\gc} \defeq (\pure,
  H^\mathrm{code,\square} \; \sigrets \; \sigcloss \; \mscode \; \gc,
  H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc)
\]
It is a $\pure$ region which makes it persistent in future worlds.
The memory relation $H^\mathrm{code,\square} \; (\dots)$ is very selective and only accepts the pair $(\mscode,\mscode)$ and only on the following conditions: The code memory $\mscode$ must be well-formed according to $\vdash$.
Further, all the words contained in $\mscode$ must be non-linear and in the value relation, but as we will see later we have two value relations, a trusted and an untrusted, and the value relation used depends on whether $\mscode$ is trusted.
For now it suffices to know that the untrusted value relation contains everything that untrusted code can safely have.
The trusted value relation includes everything that is in the untrusted value relation, but also the things that are not safe to pass away in a call, e.g.\ return seals.
If the addresses of $\mscode$ are in $\ta$, then the trusted value relation is used, and if $\mscode$ is disjoint with $\ta$, then the untrusted value relation is used.
Further, if $\ta$ is untrusted, then $\sigrets$ has to be empty which means that $\mscode$ cannot contain return seals.

@@

\begin{multline*}
  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; (\ta,\stkb) \; \sigma \; \hat{W}= \\
  \begin{array}[t]{l}
\left\{
    \begin{array}{l}
      \npair{\left(\retptrc(\baddr,\eaddr,\aaddr'+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \mid 
        \sigrets \subseteq \gsigrets \tand \\
        \quad\dom(\mscode) \subseteq \ta \tand 
        \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
        \quad\aaddr = \aaddr' + \retoffset \tand \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand\\
        \quad \sigma = \sigma_b + \offsigma \in \sigrets \tand
        \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
    \end{array}
      \right\} \cup \\
\left\{
    \begin{array}{l}
      \npair{(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1))} \mid
        \sigrets \subseteq \gsigrets \tand 
        \dom(\mscode) \subseteq \ta \tand\\
        \quad \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]},\ms_S,\ms_T \ldotp\\
        \qquad\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
        \qquad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
        \qquad \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
        \qquad \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
    \end{array}
    \right\} \\
  \end{array}\\
    \text{ if } \sigma \in \sigrets\\
\end{multline*}
We leave out the formal definition of $H^\mathrm{code,\square}_\sigma \; (\dots)$ for $\sigma \in \sigcloss$ and provide an informal description.
What closure seals can be used for depends on whether the code is trusted.
If the code is untrusted, i.e.\ $\dom(\mscode) \# \ta$, then the seal can be used for everything in the untrusted value relation can be sealed which correspond to everything the code has access to.
On the other hand, if the code is trusted, then it depends on whether the sealable in question is executable.
If it is executable, then it can be in the trusted value relation because we want to allow sealing of capabilities for the trusted code; but if it isn't executable, then the sealable has to be in the untrusted value relation.

>>>>

\subsubsection{Code region}
The code region is a central region to the logical relation used to model the code memory of a component.
It is especially interesting because it specifies what the return and closure seals in a component can be used to seal.
The code region is defined as follows:
\[
  \codereg{\sigrets,\sigcloss,\mscode,\gc} \defeq (\pure,
  H^\mathrm{code,\square} \; \sigrets \; \sigcloss \; \mscode \; \gc,
  H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc)
\]
The code region is $\pure$ which makes it persistent in future worlds and reflects the fact that it is addressed by normal capabilities.
The memory relation $H^\mathrm{code,\square} \; (\dots)$ is very selective and only accepts the pair $(\mscode,\mscode)$ and only on the following conditions: The code memory $\mscode$ must be well-formed according to $\vdash$.
Further, all the words contained in $\mscode$ must be non-linear and in the value relation, but as we will see later we have two value relations, a trusted and an untrusted, and the value relation used depends on whether $\mscode$ is trusted.
For now it suffices to know that the untrusted value relation contains everything that untrusted code can safely have.
The trusted value relation includes everything that is in the untrusted value relation, but also the things that are not safe to pass away in a call, e.g.\ return seals.
If the addresses of $\mscode$ are in $\ta$, then the trusted value relation is used, and if $\mscode$ is disjoint with $\ta$, then the untrusted value relation is used.
Further, if $\ta$ is untrusted, then $\sigrets$ has to be empty which means that $\mscode$ cannot contain return seals.

@@

the seal interpretation function $h^\mathrm{code,\square}_\sigma \; (\dots)$ in the code region dictates the return seals and closure seals can be used to seal.
the return seals can only be used to seal return pointers, but only when they actually represent meaning full returns.
for instance, a code return pointer should actually be associated with a trusted call, i.e.\ it should point to the address just after a call.
the corresponding capability on \trgcm{} should point to the address just after the \texttt{xjmp} in the call code (figure~\ref{fig:call-code}) as it needs to do a couple of things to complete the call safely.
further, the call they are associated with has to be the call that uses the seal in question.
the code return pointer is not linear which means that it can be reused which is fine as it can only be used to return from the same call.
on the other hand, the data return pointers are linear which is what prevents reuse of a return pair.
multiple nested calls from the same trusted call will result in multiple stack frames that are each sealed with the same return seal.
the seal interpretation function should allow a data return pointer for any of those pointers, so it simply requires that there exists a region $\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}$ (a region that only accepts $(\ms_S,\ms_T)$ and requires the contents to be in the untrusted value relation) in the part of the world that governs the call stack.
The data return pointer should point to that stack frame.
A region in the world for the call stack is also associated with a return address, and it should be the return address of the call that uses the return seal in question.
\todo[inline]{Consider writing something about how this relates to prove? We have to show that there is a stack frame, but we can rely on it when we receive a capability with the correct seal.}
  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; (\ta,\stkb,\gsigrets,\gsigcloss) \; \sigma \; \hat{W}= \\
  \begin{array}[t]{l}
      \npair{\left(\retptrc(\baddr,\eaddr,\aaddr'+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \mid 
        \sigrets \subseteq \gsigrets \tand \\
        \quad\dom(\mscode) \subseteq \ta \tand 
        \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
        \quad\aaddr = \aaddr' + \retoffset \tand \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand\\
        \quad \sigma = \sigma_b + \offsigma \in \sigrets \tand
        \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
      \right\} \cup \\
      \npair{(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1))} \mid
        \sigrets \subseteq \gsigrets \tand 
        \dom(\mscode) \subseteq \ta \tand\\
        \quad \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]},\ms_S,\ms_T \ldotp\\
        \qquad\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
        \qquad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
        \qquad \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
        \qquad \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
    \right\} \\
  \end{array}\\
    \text{ if } \sigma \in \sigrets
We leave out the formal definition of $H^\mathrm{code,\square}_\sigma \; (\dots)$ for $\sigma \in \sigcloss$ and provide an informal description.
What closure seals can be used for depends on whether the code is trusted.
If the code is untrusted, i.e.\ $\dom(\mscode) \# \ta$, then the seal can be used for everything in the untrusted value relation can be sealed which correspond to everything the code has access to.
On the other hand, if the code is trusted, then it depends on whether the sealable in question is executable.
If it is executable, then it can be in the trusted value relation because we want to allow sealing of capabilities for the trusted code; but if it isn't executable, then the sealable has to be in the untrusted value relation.
